Incremental Forward Reasoning for White-Box~Proof~Search

### 2025-02-18 paper.tex
Move example rules states into a figure

For slot 2, unification yields the substitution $σ ≔ \{y ↦ 301, z ↦ 302\}$, so we insert $h₁$ into $H_{y}(2, 301)$.
No matches at level 1 can be extended by $h₁$ since
\[
  \bigcap_{x \in \vars(2)} M_{x}(1, σ(x)) = M_{y}(1, 301) = \emptyset
\]
At this point, $H$ and $M$ are in the following state.
\begin{center}
\begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
  \node[xshift = 0] {$H_{y}$}
    child {node {2}
      child {node {$301$}
        child {node {$\{h₁\}$}}}
    %  child {node {$300$}
    %    child {node {$\{\}$}}}
    };
  \node[xshift=160] {$M_y$}
    child {node {1}
      child {node {$302$}
        child {node {$\{[h₁]\}$}}}
    %  child {node {$301$}
    %    child {node {$\{\}$}}}
    };
    %child {node[xshift=20] {2}
    %  child {node {$301$}
    %    child {node[yshift=-2em] {$\{\}$}}}};
\end{tikzpicture}
\end{center}

For slot 2, we add $h₂$ to $H_{y}(2, 300)$.
There are no compatible matches at level 1, so we are done with $h₂$, yielding the following final rule state.
\begin{center}
    \begin{tikzpicture}[outer sep=auto, level distance=12mm, scale=0.50]
      \node[xshift = 0] {$H_y$}
        child {node {2}
          child {node {$301$}
            child {node {$\{h₁\}$}}}
          child {node {$300$}
            child {node {$\{h₂\}$}}}
        };
      \node[xshift=160] {$M_y$}
        child {node {1}
          child[xshift=-20] {node {$302$}
            child {node {$\{[h₁]\}$}}}
          child[xshift=20] {node {$301$}
            child {node {$\{[h₂]\}$}}}
        };
    \end{tikzpicture}
    \end{center}

\paragraph{Example 2.}
We now consider a rule $r₂ : \All{x~y}{A~x~y → B~x~y → C~x~y}$ with multiple shared variables.
We add hypotheses $h₁ : B~a~a$, $h₂ : A~a~b$ and $h₃ : B~a~b$ to the rule state of $r₂$, starting with $h₁$:
% NOTE : remove comments to get full trees.
\begin{center}
  \begin{tikzpicture}[level distance=12mm, scale=0.50]
    \node[xshift = 30] {$H_x$}
        %child[xshift = -20] {node {1}
        %    child {node {$\{ \}$}}}}

@@

@@

    \node[xshift = 110] {$H_y$}
        %child[xshift = -20] {node {1}
        %        child {node {$\{ \}$}}}}

@@

@@

    \node[xshift = 190] {$M_x$};
        %child[xshift = -20] {node {1}
        %        child {node {$\{ \}$}}}}
    \node[xshift = 270] {$M_y$};
        %child[xshift = -20] {node {1}
        %        child {node {$\{ \}$}}}}

@@

\end{center}
At this point there are no matches, so there is nothing to extend.
We now add $h_2 : A~a~b$, which initially generates a match $m = [h₂]$ that is added to $M$:
% NOTE : uncomment to get full trees.
\begin{center}

@@

@@

\end{center}
To find hypotheses compatible with $m$, we consider the set of variables $\vars(\lvl(m) + 1)$.
These are the variables shared between slot $\lvl(m) + 1 = 2$ and all previous slots, i.e.\ $x$ and $y$.
Hence the compatible hypotheses are
\[
  \bigcap_{x \in \vars(\lvl(m) + 1)} H_{x}(i + 1, \sub(m)(x)) = H_{x}(2, a) \cap H_{y}(2, b) = \emptyset
\]
This means the match cannot currently be extended.
We now turn to the remaining hypothesis $h_3 : B~a~b$, adding it first to $H$:
\begin{center}

@@

@@

\end{center}
Since $h₃$ matches slot $i = 2$, the matches that can be extended by $h₃$ are
\[
  \bigcap_{x \in \vars(i)} M_{x}(i - 1, \sub_{i}(x)) = M_{x}(1, a) \cap M_{y}(1, b) = \{[h₂]\}
\]
We have found the complete match $[h₂, h₃]$ and can add a new hypothesis of type $C~a~b$.

>>>>

For slot 2, unification yields the substitution $σ ≔ \{y ↦ 301, z ↦ 302\}$, so we insert $h₁$ into $H_{y}(2, 301)$.
No matches at level 1 can be extended by $h₁$ since
\[
  \bigcap_{x \in \vars(2)} M_{x}(1, σ(x)) = M_{y}(1, 301) = \emptyset
\]
At this point, $H$ and $M$ are in the state shown in Fig.~\ref{fig:example1:state1}.

\begin{figure}
  \centering
  \begin{minipage}{\textwidth}
    \begin{subcaptiongroup}
      \subcaptionlistentry{}
      \begin{minipage}{0.5\textwidth}
          \node at (-2,0) {\captiontext*{}\label{fig:example1:state1}};
          \node[xshift = 0] {$H_{y}$}

@@

              child {node {$301$}
                child {node {$\{h₁\}$}}}
            %  child {node {$300$}

@@

          \node[xshift=80] {$M_y$}
              child {node {$302$}
                child {node {$\{[h₁]\}$}}}
            %  child {node {$301$}
            %child {node[xshift=20] {2}
            %  child {node {$301$}
            %    child {node[yshift=-2em] {$\{\}$}}}};
      \end{minipage}
    \end{subcaptiongroup}
    \begin{subcaptiongroup}
      \subcaptionlistentry{}
      \begin{minipage}{0.5\textwidth}
          \node at (-2,0) {\captiontext*{}\label{fig:example1:state2}};
          \node[xshift = 0] {$H_y$}

@@

              child {node {$301$}
                child {node {$\{h₁\}$}}}
              child {node {$300$}
                child {node {$\{h₂\}$}}}

@@

          \node[xshift=80] {$M_y$}
              child[xshift=-20] {node {$302$}
                child {node {$\{[h₁]\}$}}}
              child[xshift=20] {node {$301$}
                child {node {$\{[h₂]\}$}}}

@@

      \end{minipage}
    \end{subcaptiongroup}
  \end{minipage}
  \\[1em]
  \begin{subcaptiongroup}
    \subcaptionlistentry{}
    \begin{minipage}{\textwidth}

@@

        \node at (-2,0) {\captiontext*{}\label{fig:example2:state1}};

@@

    \end{minipage}
  \end{subcaptiongroup}
  \\[1em]
  \begin{subcaptiongroup}
    \subcaptionlistentry{}
    \begin{minipage}{\textwidth}

@@

        \node at (-2,0) {\captiontext*{}\label{fig:example2:state2}};

@@

    \end{minipage}
  \end{subcaptiongroup}
  \caption{Rule states encountered during Examples~1 and~2}
\end{figure}

For slot 2, we add $h₂$ to $H_{y}(2, 300)$.
There are no compatible matches at level 1, so we are done with $h₂$, yielding the final rule state in Fig.~\ref{fig:example1:state2}.

\paragraph{Example 2.}
We now consider a rule $r₂ : \All{x~y}{A~x~y → B~x~y → C~x~y}$ with multiple shared variables.
We add hypotheses $h₁ : B~a~a$, $h₂ : A~a~b$ and $h₃ : B~a~b$ to the rule state of $r₂$, starting with $h₁$, which is added to $H$ with slot 2.
There are no matches yet, so there is nothing to extend.
We then add $h_2 : A~a~b$, which initially generates a match $m = [h₂]$ that is added to $M$, yielding the rule state in Fig.~\ref{fig:example2:state1}.
% NOTE : uncomment to get full trees.
To find hypotheses compatible with $m$, we consider the set of variables $\vars(\lvl(m) + 1)$.
These are the variables shared between slot $\lvl(m) + 1 = 2$ and all previous slots, i.e.\ $x$ and $y$.
Hence the compatible hypotheses are
\[
  \bigcap_{x \in \vars(\lvl(m) + 1)} H_{x}(i + 1, \sub(m)(x)) = H_{x}(2, a) \cap H_{y}(2, b) = \emptyset
\]
This means the match cannot currently be extended.
We now turn to the remaining hypothesis $h_3 : B~a~b$, adding it first to $H$.
This yields the rule state in Fig.~\ref{fig:example2:state2}.
Since $h₃$ matches slot $i = 2$, the matches that can be extended by $h₃$ are
\[
  \bigcap_{x \in \vars(i)} M_{x}(i - 1, \sub_{i}(x)) = M_{x}(1, a) \cap M_{y}(1, b) = \{[h₂]\}
\]
This produces the complete match $[h₂, h₃]$, so we can add a new hypothesis of type $C~a~b$.
